---
title: "Exercise 3 - ORGB672"
author: "Xénia Sozonoff"
date: "2023-03-25"
output:
  pdf_document: default
  html_document: default
---

## Load data

```{r}
library(arrow)
library(tidyverse)
applications <- read_parquet("~/Downloads/app_data_sample.parquet")
edges <- read_csv("~/Downloads/edges_sample.csv")
```

```{r}
applications
```
```{r}
edges
```
## Get gender for examiners 

```{r}
library(gender)

# Get a list of first names without repetitions
examiner_names <- applications %>% 
  distinct(examiner_name_first)

examiner_names
```

```{r}
# Get a table of names and gender
library(tidyverse)
examiner_names_gender <- examiner_names %>% 
  do(results = gender(.$examiner_name_first, method = "ssa")) %>% 
  unnest(cols = c(results), keep_empty = TRUE) %>% 
  select(
    examiner_name_first = name,
    gender,
    proportion_female
  )

examiner_names_gender
```

```{r}
# Remove extra colums from the gender table
examiner_names_gender <- examiner_names_gender %>% 
  select(examiner_name_first, gender)

# Joining gender back to the dataset
applications <- applications %>% 
  left_join(examiner_names_gender, by = "examiner_name_first")

# Cleaning up
rm(examiner_names)
rm(examiner_names_gender)
gc()
```

## Guess the examiner’s race

```{r}
library(wru)

examiner_surnames <- applications %>% 
  select(surname = examiner_name_last) %>% 
  distinct()

examiner_surnames
```

```{r}
examiner_race <- predict_race(voter.file = examiner_surnames, surname.only = T) %>% 
  as_tibble()
```

```{r}
examiner_race
```
```{r}
examiner_race <- examiner_race %>% 
  mutate(max_race_p = pmax(pred.asi, pred.bla, pred.his, pred.oth, pred.whi)) %>% 
  mutate(race = case_when(
    max_race_p == pred.asi ~ "Asian",
    max_race_p == pred.bla ~ "black",
    max_race_p == pred.his ~ "Hispanic",
    max_race_p == pred.oth ~ "other",
    max_race_p == pred.whi ~ "white",
    TRUE ~ NA_character_
  ))

examiner_race
```
```{r}
# Removing extra columns
examiner_race <- examiner_race %>% 
  select(surname,race)

applications <- applications %>% 
  left_join(examiner_race, by = c("examiner_name_last" = "surname"))

rm(examiner_race)
rm(examiner_surnames)
gc()
```

## Examiner’s tenure

```{r}
library(lubridate) # to work with dates

examiner_dates <- applications %>% 
  select(examiner_id, filing_date, appl_status_date) 

examiner_dates
```

```{r}
examiner_dates <- examiner_dates %>% 
  mutate(start_date = ymd(filing_date), end_date = as_date(dmy_hms(appl_status_date)))
```

```{r}
examiner_dates <- examiner_dates %>% 
  group_by(examiner_id) %>% 
  summarise(
    earliest_date = min(start_date, na.rm = TRUE), 
    latest_date = max(end_date, na.rm = TRUE),
    tenure_days = interval(earliest_date, latest_date) %/% days(1)
    ) %>% 
  filter(year(latest_date)<2018)

examiner_dates
```

```{r}
applications <- applications %>% 
  left_join(examiner_dates, by = "examiner_id")

rm(examiner_dates)
gc()
```

##  Pick two workgroups to focus on 

```{r}
# Filter for workgroups 174 and 165
workgroup1 <- applications %>% 
  filter(str_sub(examiner_art_unit, 1, 3) == "174")
workgroup2 <- applications %>% 
  filter(str_sub(examiner_art_unit, 1, 3) == "165")

# Remove the rows with NA values in the Gender, Race and Tenure columns
workgroup1 <- workgroup1[!is.na(workgroup1$gender) & !is.na(workgroup1$race)  & !is.na(workgroup1$tenure_days), ]
workgroup2 <- workgroup2[!is.na(workgroup2$gender) & !is.na(workgroup2$race)  & !is.na(workgroup2$tenure_days), ]

# Calculate proportion of examiners by gender and race
prop_gender <- data.frame(
  Workgroup = c("174", "165"),
  Proportion_Male = c(mean(workgroup1$gender == "male"), mean(workgroup2$gender == "male")),
  Proportion_Female = c(mean(workgroup1$gender == "female"), mean(workgroup2$gender == "female"))
)

prop_race <- data.frame(
  Workgroup = c("174", "165"),
  Proportion_White = c(mean(workgroup1$race == "white"), mean(workgroup2$race == "white")),
  Proportion_Black = c(mean(workgroup1$race == "black"), mean(workgroup2$race == "black")),
  Proportion_Asian = c(mean(workgroup1$race == "Asian"), mean(workgroup2$race == "Asian")),
  Proportion_Hispanic = c(mean(workgroup1$race == "Hispanic"), mean(workgroup2$race == "Hispanic")),
  Proportion_Other = c(mean(workgroup1$race == "other"), mean(workgroup2$race == "other"))
)

# Calculate mean tenure of examiners
mean_tenure <- data.frame(
  Workgroup = c("174", "165"),
  Mean_Tenure = c(mean(workgroup1$tenure_days), mean(workgroup2$tenure_days))
)

# Combine results into summary statistics table
summary_statistics <- merge(prop_gender, prop_race, by = "Workgroup")
summary_statistics <- merge(summary_statistics, mean_tenure, by = "Workgroup")

summary_statistics
```

```{r}
# Plot the gender comparison of the two workgroups
library(ggplot2)

ggplot() +
  geom_bar(data = workgroup1, aes(x = gender, fill = "174"), alpha = 0.7) +
  geom_bar(data = workgroup2, aes(x = gender, fill = "165"), alpha = 0.7) +
  scale_fill_manual(name = "Workgroup", values = c("174" = "coral", "165" = "skyblue")) +
  labs(title = "Gender Comparison Between Workgroups", x = "Gender", y = "Number of Examiners") +
  theme_classic()
```

```{r}
ggplot() +
  geom_bar(data = workgroup1, aes(x = race, fill = "174"), alpha = 0.7) +
  geom_bar(data = workgroup2, aes(x = race, fill = "165"), alpha = 0.7) +
  scale_fill_manual(name = "Workgroup", values = c("174" = "coral", "165" = "skyblue")) +
  labs(title = "Race Comparison Between Workgroups", x = "Race", y = "Number of Examiners") +
  theme_classic()
```

```{r}
# Plot the tenure comparison of the two workgroups
ggplot() +
  geom_density(data = workgroup1, aes(x = tenure_days, color = "174"), alpha = 0.7) +
  geom_density(data = workgroup2, aes(x = tenure_days, color = "165"), alpha = 0.7) +
  scale_color_manual(name = "Workgroup", values = c("174" = "coral", "165" = "skyblue")) +
  labs(title = "Tenure Comparison Between Workgroups", x = "Tenure days", y = "Density") +
  theme_classic()
```

There are noticeable differences in the gender and race composition of workgroup1 (174) and workgroup2 (165). Workgroup1 has a higher proportion of male and white examiners, while workgroup2 has a higher proportion of female, black, Asian, and Hispanic examiners. Both workgroups have a predominance of white examiners over other races. The difference in mean tenure between the two workgroups is small, with workgroup2 having a slightly higher mean tenure.

## 3 Centrality scores 

```{r}
# Filter edges to include only the selected workgroups
workgroup_edges <- edges%>%
  filter(ego_examiner_id %in% workgroup1$examiner_id | ego_examiner_id %in% workgroup2$examiner_id |
           alter_examiner_id %in% workgroup1$examiner_id | alter_examiner_id %in% workgroup2$examiner_id)

applications <- applications %>%
  filter(!is.na(gender), !is.na(race), !is.na(tenure_days))

# Create the advice network
advice_network <- graph_from_data_frame(workgroup_edges, directed = TRUE)

#Check if the advice network is connected 
is.connected(advice_network)

# Get the clusters in the network
advice_clusters <- clusters(advice_network)

# Get the nodes in the largest cluster
largest_cluster <- which.max(advice_clusters$csize)
cluster_nodes <- as.numeric(membership(advice_clusters)[which(advice_clusters$membership == largest_cluster)])

# Remove isolated nodes
isolated_nodes <- V(advice_network)[degree(advice_network, mode="total") == 0]
advice_network <- delete.vertices(advice_network, isolated_nodes)

# Extract the largest connected component
largest_component <- components(advice_network)$membership == which.max(components(advice_network)$csize)
advice_network <- induced.subgraph(advice_network, largest_component)

# Calculate centrality scores
centrality_scores <- data.frame(
  name = as.character(V(advice_network)$name),
  degree = degree(advice_network, mode = "in"),
  closeness = closeness(advice_network, mode = "in"),
  betweenness = betweenness(advice_network, directed = TRUE)
)

# Print the top 10 examiners by degree centrality
top_degree_centrality <- centrality_scores %>% 
  arrange(desc(degree)) %>% 
  head(10)

top_degree_centrality

# Print the top 10 examiners by closeness centrality
top_closeness_centrality <- centrality_scores %>% 
  arrange(desc(closeness)) %>% 
  head(10)

top_closeness_centrality

# Print the top 10 examiners by betweenness centrality
top_betweenness_centrality <- centrality_scores %>% 
  arrange(desc(betweenness)) %>% 
  head(10)

top_betweenness_centrality
```

### Characterize and discuss the relationship between centrality and other examiners’ characteristics

It is difficult here to conclude and to draw a clear relationship between centrality and other examiners' characteristics. However, we can see that the examiners with the highest degree centrality do not necessarily have the highest closeness centrality scores.There was certainly a gap in the code, I should have created tables that allowed to understand the relationship between centrality scores and gender, race, and tenure days. I should have added this information as well as which workgroup I was referring to, in order to draw a correct conclusion.